---
title: "POL304 Assignment 1"
author: "Talia Fabregas"
date: "23 January 2024"
format: pdf
editor: visual
---

```{r}
## loading packages
# install.packages("tidyverse")
# install.packages("janitor")
# install.packages("dplyr")
library(tidyverse)
library(janitor)
library(dplyr)

## load the data
chechen_data <- read_csv("chechen.csv")
```

## Question 1

What is the research question? (Remember that a research question is general, it does not name specific places or policy areas. Also note that a research question must be open-ended, it is not a yes/no question).

The research question that this paper investigates is: "Does a state's use of indiscriminate violence incite insurgent attacks?"

The research question that this paper investigates is the extent to which a state's use of indiscriminate violence incites insurgent attacks.

## Question 2

What is the theory used in the article? \[Word limit: 150 words\] Do not simply copy/paste from the article. Instead, think carefully what is the set of logically consistent assumptions that lead to the article\'s main prediction. Focus on the broad strokes rather than the details. Ask yourself what are the most basic relationships/conditions that must hold in order to derive the article\'s main prediction. List these assumptions and the relationship among them. For example, if the first assumption is that the Earth is flat and the second assumption is that the edge of the Earth is in the ocean, then a logically consistent prediction would be that we should observe water pouring off of the edge of the Earth into the void. Notice that neither the assumptions nor the prediction need to be true. The assumptions are simplifications of the world; the only requirement is that they are logically consistent with one another.

The theory used in this article is ...

## Question 3

What is the central research hypothesis. \[Answer in one sentence\] Remember that, just like the research question and the theory, the hypothesis must be stated in general, rather than specific, terms.

The central research hypothesis is:

## Question 4

75 villages were shelled by Russia, and 50 were not

```{r}
# count how many unique villages there are
num_villages <- length(unique(chechen_data$village))

# subset to only include the villages that were shelled
shelled_villages <- chechen_data |>
  filter(fire == 1)

# count the number of unique villages that were shelled
num_shelled = length(unique(shelled_villages$village))

# get the number of unique villages that were not shelled
num_not_shelled = num_villages - num_shelled

# create a data frame
q4 <- data.frame(num_villages_shelled = num_shelled,
                num_villages_not_shelled = num_not_shelled)

# print the data frame
print(q4)
```

## Question 5

```{r}
# shelled villages

# non-shelled villages
```

## Question 6

## Question 7

## Question 8

## Question 9

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
